---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# OvertureMaps R Package

This package provides functions to interact with OvertureMaps datasets. It includes utilities for fetching data from S3 and converting it to sf objects for spatial analysis.

<!-- badges: start -->

[![R-CMD-check](https://github.com/denironyx/overturemapsr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/denironyx/overturemapsr/actions/workflows/R-CMD-check.yaml)

<!-- badges: end -->

Overture Maps provides free and open geospatial map data, from many different sources and normalized to a common schema. This tool helps to download Overture data within a region of interest and converts it to a few different file formats.

## Installation

To install this package, you need to have the necessary dependencies installed. You can install the package using the following commands:

``` r
# install.packages("devtools")
devtools::install_github("denironyx/overturemapsr")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(overturemapsr)
## basic example code
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.

## Try it out

Example lisbon building
```{r }
library(overturemapsr)


```

## Functions

- `get_all_overture_types`:  This function returns all available OvertureMaps theme types. 
- `dataset_path`: This function returns the S3 path for the specified Overture dataset type. 
- `record_batch_reader`: This function retrieves a filtered dataset from the specified Overture dataset type, optionally within a bounding box, and converts it to an sf object.

#### Parameters

-   overture_type: Character. Required. The type of feature to select. Examples include 'building', 'place', etc. To learn more, run get_all_overture_types().
-   bbox: Numeric vector. Optional. A bounding box specified as c(xmin, ymin, xmax, ymax). It is recommended to use a bounding box to limit the dataset size and processing time. Without a bounding box, processing the entire dataset (e.g., buildings over 2 billion) can be time-consuming.


[Bounding Box Tool from Klokantech](https://boundingbox.klokantech.com/): Use this tool to find and specify bounding boxes of interest for your data queries.

### Notes
This package was inspired and based on the [OvertureMaps Python Repository](https://github.com/OvertureMaps/overturemaps-py/tree/main). The functionality are similar and I will ensure to make the package stay updated.


##### Acknowledgement
Another inspiration for this package was based on the blog post by [Kyle Walker](https://walker-data.com/posts/overture-buildings/). Kudos!!!
